\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-06-16 eg6-letters/NIM/EG6_RTPC_NIM.tex
Update to elsevier framework
Comments from Sebastian

\documentclass[twocolumn,showpacs,superscriptaddress,groupedaddress]{revtex4}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}  
\usepackage{amssymb}  
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue, citecolor=blue}

\begin{document}

@@

Thomas Jefferson National Accelerator Facility (TJNAF), in Newport News, Virginia,
USA, provides high power electron beams of energy up to 6 GeV and 100$\%$ duty cycle
to three experimental Halls (A, B, C) simultaneously. The CLAS 
spectrometer~\cite{CLASref}, located
in Hall-B, is composed of several sub-detectors and two magnets. 
Figure~\ref{fig:CLAS} shows a three dimensional 
representation of the baseline CLAS spectrometer:
\begin{itemize}
 \item Three regions of Drift Chambers (DC) for the tracking of charged 
    particles~\cite{DCref} (in blue on the figure).
 \item Superconducting toroidal magnet (in yellow) to bend the trajectories 
    of charged particles, thus allowing momentum measurement with the DC tracking information.
 \item Cerenkov Counters (CC) to distinguish electrons from other negative 
    particles, such as pions, at lower energies~\cite{CCref} (in pink).
 \item Scintillation Counters (SC) to identify hadrons by measuring their 
    time of flight (TOF)~\cite{TOFref} (in red).
 \item Electromagnetic Calorimeters (EC) to distinguish electrons from other negative 
    particles at higher energies and to detect neutral particles (neutrons and 
    photons in particular)~\cite{ECref} (in green).
\end{itemize}

@@

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a hadron) necessitates an upgrade
of this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles, where the acceptance of the EC is poor. To extend the detection range,
an inner calorimeter (IC) was built for the E01-113 experiment in 2005.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}. To protect this detector from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the IC. 
In addition, the recoil nuclei of the coherent DVCS on helium was detected 
using a RTPC developed to track low energy nuclear fragments. The solenoid field
serving as the analyzing magnet for the particle tracking in the RTPC. This 
setup was used during a three months experimental run~\cite{proposal1,proposal2}
in 2009 with a longitudinally polarized electron beam of 130~nA 
and energy of 6.064 GeV incident on a gaseous $^{4}$He target.

@@

\section{CLAS RTPC design} \label{sec_design}

@@

The CLAS RTPC is a 250~mm long cylinder of 158~mm diameter, leaving just enough 
room to fit pre-amplifiers between the RTPC outer shell and the solenoid. The 
electromagnetic field is directed perpendicularly to the beam direction, 
such that drifting electrons are pushed away from the beam line. These electrons 
are amplified by three layers of curved GEMs and detected by the readout system 
on the external shell of the detector as illustrated in the 
Fig.~\ref{fig:RTPC_1_4}. The CLAS RTPC is segmented into two halves with 
independent GEM amplification systems that cover about 80\% of the azimuthal 
angle.

We detail here the different regions shown in Fig.~\ref{fig:RTPC_1_4} starting 
from the beam line towards larger radius:\\
\begin{itemize}
   \item First, the 6 atm helium-4 target extends along the beam line axis, it 
      is 284~mm long with a 3~mm radius wall made of a 27-$\mu$m-thick Kapton.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      that is electronically grounded.
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode and is set at 4260~V.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first GEM, 60 mm away 
      from the beam axis. The average electric field in this region is about 550~V/cm 
      and perpendicular to the beam.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied to each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Pre-amplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

The drift gas used in the experiment is a 80-20\% Ne-DME mixture, this choice 
has been made in order to balance the energy deposit with a reasonable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the perpendicular 5~T field, we would have a Lorentz angle of 
about 23$^\circ$ with this gas mixture. This is already a significant angle,
a heavier gas would have larger angle increasing significantly
the length of the drift paths as the electron would 
make a long spiral around the beam line to reach the GEMs.

One of the important challenge in developing the radial TPC was how to obtain a 
good support structure of the GEM foils to allow a tractable but highly precise 
installation of the GEMs. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfully realized these prerequisites 
by using fiber glass rings glued to each ends of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations. The rigidity of the GEM foils was enough
for the structure to be self-supporting and only the backward part of the 
cylinder was fixed to the main mechanical structure. This design only left a 
light fiberglass ring in the forward region, reducing to a minimum secondary 
interactions.

@@

The RTPC electron collection system has 3200 readout pads. These elements are
located at the end of the amplification region, 69 mm from the central axis.
Figure~\ref{fig:PADs} shows a schematic drawing of the size and 
configuration of the pads. Each readout pad is 5 mm in length and 4.45 mm in 
width. The shift between the rows permits the reduction of aliasing. Each half of the 
RTPC has 40 rows and 40 columns of pads. The shaded region in Fig.~\ref{fig:PADs} 
shows how pads are grouped to 16 channels pre-amplifiers. The pre-amplifier boards, 
already employed in the BoNuS RTPC~\cite{BONUS-NIM}, serve the dual purpose of 
inverting the RTPC signals polarity -- from negative to positive -- to match the 
requirements of the subsequent readout system, and driv the 6~m long ribbon 
cable that connects to it.

The readout system is based on the Front End Electronic (FEE) boards originally 
developed for the ALICE TPC readout system at CERN~\cite{ALICE-FEE}. Each 
readout channel is made of three main components: a charge sensitive, low 
impedance shaping amplifier, a 10 bit, 25-MHz digitizer, and a digital circuit 
that implements online-processing algorithms (pedestal subtraction, zero 
suppression, and tail cancellation). The amplifier is implemented in a 
fully-analogue ASIC, PASA~\cite{ALICE-PASA}, while the subsequent signal 
digitization and elaboration part is implemented in a digital ASIC, 
ALTRO~\cite{ALICE-ALTRO}. Both ASICs host 16 channels. The full readout 
chain is implemented on a single Front End Card (FEC), hosting 128 channels. 
FECs are hosted in custom crates, mounted close to the detector. A readout 
control unit (RCU) board is used to distribute the trigger signal to FECs 
and for data readout. Each RCU can handle up to 25 FECs with communication 
between them performed trough a 
custom back-plane, implementing a low-voltage signal bus. The RCU communicates 
with the CLAS data acquisition (DAQ) system 
trough a 200 MB/s optical link, connected to a data acquisition PC hosting 
a ReadOut Receiver Card (RORC) used to interface with 
the CLAS DAQ system. An Ethernet link is also present, for slow-controls and 
monitoring.

Triggering and data-readout scheme is as follows. When a Level-1 (L1) trigger 
is received by a FEC-board, a programmable number of samples $N_0$ is digitized 
and stored in a data memory, and hence processed by the digital circuit. Upon 
arrival of Level-2 trigger, the latest processed samples are saved in a 
multi-event memory, individual to each-channel, ready for readout by the RCU.  
In our setup, the RCU receives the trigger signal from the CLAS Trigger 
Supervisor system, and distributes both L1 and L2 signals to all the FECs in 
the crate, with a definite L1-L2 trigger latency. Such latency is programmable, 
and has to be long enough to accommodate the digitization of the $N_0$ samples 
per channel.  In the current implementation, $N_0=100$ samples/channel are 
acquired for each trigger, operating ALTRO at the reduced sampling frequency of 
$10$~MHz. The L1-L2 latency is thus fixed to 15 $\mu$s.

The trigger signal also initiates RCU readout operation from FEC boards. All 
the measured samples from active channels are reported, together with a channel 
identifier and a time-stamp, to the ROC application, that in turns sends them 
on the main event builder. During data reconstruction, the acquired samples are 
processed to obtain, for each readout pad, the accumulated charge (ADC) and the 
pulse time (T). Since pulse time was obtained as the time-stamp of the first 
sample above threshold, the resolution is equivalent to the ALTRO sampling 
time of 100~ns.

\begin{figure}[tb]
   \centering
   \includegraphics[width=.45\textwidth]{fig/ReadoutScheme.png}
   \caption[]{Schematic representation of the RTPC readout system, showing} 
   \label{fig:ReadoutScheme}
\end{figure}

@@

In order to read all the detector readout pads, four FEC crates are used, 
each equipped with 6 boards, plus a ROC. A schematic of the readout system, 
for a single crate, is reported in Fig.~\ref{fig:ReadoutScheme}. This 
configuration permits to reduce the dead time associated to FEC readout 
operations and to scale linearly with the number of boards in the crate. During the 
2009 run, the system was successfully operated with a DAQ rate of 3.1 kHz, with 
a live time of $70 \%$.

@@

The timing information is used to infer the origin of the charge and 
then the trajectory of the detected particle
resulting in momentum measurement. Going from a collection of 
times to a momentum measurement requires a good knowledge of the drift speed 
and drift paths followed by the electrons released in the gas. The recorded ADCs give 
the deposited energy per unit of length ($\small{\frac{dE}{dX}}$) which, 
together with the momentum calculated from the trajectory, enable the particle 
identification.

In this section we will detail the methods used to calibrate the drift speed,
drift paths and gains of the detector. Drift speed and paths were initially
calculated using the MAGBOLTZ~\cite{MAGBOLTZ} program, then refined using
data to account for variations of the run conditions. We always assume 
cylindrical symmetry in the chamber for the calibration, such that none of
the parameters depend on the azimuthal angle $\phi$. The initial MAGBOLTZ
calibration was improved through several iterations of the
process described below, each iteration increasing the number of events 
properly reconstructed in the RTPC. The figures presented in this section
are the one obtained while performing the last iteration of this long 
calibration process.

@@

\begin{figure}[tb]
   \centering
   \includegraphics[scale=0.42]{fig_2017/hits_time_profile.pdf}
   \caption[]{Time profile of the collected hits in one experimental run.  } 
   \label{fig:TDC_profile}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[scale=0.42]{fig_2017/RunNumber_61452_TDCmax_Zslice.pdf}
\caption{Time profile distribution for the collected hits in one experimental 
run. } \label{fig:RunNumber_61551_TDCmax_Zslice}
\end{figure}

@@

Because of the magnetic field, the drift paths are not linear in the RTPC. So 
to perform the extraction, we make a first approximation with a linear 
dependence between the radius of emission and the time of detection, and then 
refine our result. As it happens, the curvature is minimal and this process 
converge already on the second iteration.

At the end of the extraction procedure, the azimuthal difference between the detection pad and 
the ionization point ($\Delta\phi$) is extracted as a function of time. 
In Fig.~\ref{fig:DELTA_PHI_TDC}, we show the resulting data points for one 
bin, where the drift paths is easily identified and eventually fitted for 
implementation in our reconstruction codes.

@@

\begin{figure}[tb]
   \centering
   \includegraphics[scale=0.43]{fig_2017/f_dedx_p_exp_1st.pdf}
   \caption{$\small{\frac{dE}{dX}}$ vs. $p/q$ distribution for the left half of 
      the RTPC (left) and for the right half (right). Here, 
      $\small{\frac{dE}{dX}}$ is calculated using the gains of the first 
      method.  The lines are theoretical expectations from Bethe-Bloch formula 
   for $^4$He, $^3$He, $^3$H and $^2$H (d).}
\label{fig:dedx_p_exp_1st}
\end{figure}

The second method is to compare the experimental ADCs to the energy deposited in 
GEANT4 by similar simulated tracks (using the same elastic events than for the drift 
paths calibration). This requires a very good GEANT4 simulation 
including drift paths, but also the spread of the charges along the path
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the DAQ features 
that can lead to cutting out hits. After setting the simulation properly, we 
compared simulation to experiment on an event by event basis as shown in 
Fig.~\ref{fig:EVENT_adc_tdc}. In this step, the gain for each pad is calculated 
as the ratio of the measured ADCs to the simulated deposited energy.  
Then, these gains are refined using correcting factors obtained from a 
sample of good tracks. For each track, we calculate the corrected 
energy deposit on a pad and compare it to the average
deposit recorded by the other pads. 
Results of this second method are shown in Fig.~\ref{fig:dedx_p_exp_2nd},
from these we concluded that the gains of the 
second method match best the theoretical lines.

\begin{figure}[tb]
   \centering
\includegraphics[scale=0.42]{fig_2017/Event_ADC_Graph_8799283.pdf}
\caption{Simulated (upper) and experimental (lower) ADC and T distributions 
of a track. The colors indicate the pads, same color in top and bottom indicate 
that they are the same pad.}
\label{fig:EVENT_adc_tdc}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[scale=0.45]{fig_2017/f_dedx_p_exp_2nd.pdf}
\caption{$\small{\frac{dE}{dX}}$ vs. $p/q$ distribution for the left half of 
   the RTPC (left) and for the right half (right). Here, 
   $\small{\frac{dE}{dX}}$ is calculated using the gains of the second method.  
   The lines are theoretical expectations from Bethe-Bloch formula for $^4$He, 
$^3$He, $^3$H and $^2$H (d).}
\label{fig:dedx_p_exp_2nd}
\end{figure}

@@

\begin{figure}[tb]\centering
\includegraphics[scale=0.27]{fig/noisy_pad_before_rejection2.png}
\includegraphics[scale=0.27]{fig/noisy_pad_after_rejection2.png}
\caption{The ADC vs. T spectrum for an example noisy channel before (top) and 
after (bottom) noise rejection algorithms.  Only hits associated with tracks 
are included, and the selection of events and tracks is the same in both 
plots.}
\label{fig:noise}
\end{figure}

@@

\bibitem{CCref}
   G. Adams et al., The CLAS Cerenkov detector", Nucl. Inst. and Meth. A 465, 
   414 (2001).

@@

>>>>

\documentclass[preprint,5p]{elsarticle}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}  
\usepackage{amssymb}  
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue, citecolor=blue}
\usepackage[displaymath]{lineno}
\linenumbers
\begin{document}

@@

Until recently, Thomas Jefferson National Accelerator Facility (TJNAF), in 
Newport News, Virginia, USA, has provided high power electron beams of energy 
up to 6 GeV and 100$\%$ duty cycle to three experimental Halls (A, B, C) 
simultaneously. The CLAS spectrometer~\cite{CLASref}, located
in Hall-B, is composed of several sub-detectors and two magnets. 
Figure~\ref{fig:CLAS} shows a three dimensional 
representation of the baseline CLAS spectrometer:
\begin{itemize}
 \item Three regions of Drift Chambers (DC) for the tracking of charged 
    particles~\cite{DCref} (in blue on the figure).
 \item Superconducting toroidal magnet (in yellow) to bend the trajectories 
    of charged particles, thus allowing momentum measurement with the DC tracking information.
 \item Cerenkov Counters (CC) to distinguish electrons from other negative 
    particles, such as pions, at lower energies~\cite{CCref} (in pink).
 \item Scintillation Counters (SC) to identify hadrons by measuring their 
    time of flight (TOF)~\cite{TOFref} (in red).
 \item Electromagnetic Calorimeters (EC) to distinguish electrons from other negative 
    particles at higher energies and to detect neutral particles (neutrons and 
    photons in particular)~\cite{ECref} (in green).
\end{itemize}

@@

Measurement of the Deeply Virtual Compton Scattering (DVCS) process
($eH \rightarrow e' H' \gamma$, where $H$ is a nucleon or nuclei) necessitates an upgrade
of this setup.  With a 6 GeV electron beam, the majority of DVCS photons are produced
at very forward angles, where the acceptance of the EC is poor. To extend the detection range,
an inner calorimeter (IC) was built for the E01-113 experiment in 2005~\cite{FX}.
The IC is constructed from 424 lead-tungstate (PbWO$_{4}$) crystals, covering polar 
angles between 5$^{\circ}$ and 15$^{\circ}$. Each crystal is 16 cm long, corresponding
to 17 radiation lengths. The achieved energy resolution is around 3 to 4$\%$ for photon
energies between 2 GeV and 5 GeV, with an angular resolution between 3 and 5 mrad 
\cite{Hyon-suk}. To protect this detector and CLAS from the large flux of low energy M{\o}ller 
electrons, a 5~T solenoid was placed around the target to shield the IC. 
In addition, the recoil nuclei from the coherent DVCS on helium was detected 
using a RTPC developed to track low energy nuclear fragments. The solenoid field
serving as the analyzing magnet for particle tracking in the RTPC. This 
setup was used during a three months experimental run~\cite{proposal1,proposal2}
in 2009 with a longitudinally polarized electron beam of 130~nA 
and energy of 6.064 GeV incident on a gaseous $^{4}$He target.

@@

\section{RTPC design} \label{sec_design}

@@

The new CLAS RTPC is a 250~mm long cylinder of 158~mm diameter, leaving just enough 
room to fit pre-amplifiers between the RTPC outer shell and the solenoid. The 
electric field is directed perpendicularly to the beam direction, 
such that drifting electrons are pushed away from the beam line. These electrons 
are amplified by three layers of curved GEMs and detected by the readout system 
on the external shell of the detector as illustrated in the 
Fig.~\ref{fig:RTPC_1_4}. The RTPC is segmented into two halves with 
independent GEM amplification systems that cover about 80\% of the azimuthal 
angle.

We detail here the different regions shown in Fig.~\ref{fig:RTPC_1_4} starting 
from the beam line towards larger radius:\\
\begin{itemize}
   \item First, the 6 atm helium-4 target extends along the beam line axis; it 
      is 284~mm long with a 3~mm radius wall made of a 27-$\mu$m-thick Kapton.
   \item The first gas gap covers a radial range from 3 mm to 20 mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o ller electrons scattered by the beam. This 
      region is surrounded by a 4 $\mu$m thick window made of aluminized Mylar 
      that is electronically grounded.
   \item The second gap region extends between 20~mm and 30~mm and is filled with a 
      gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether (DME). This region 
      is surrounded by a 4~$\mu$m thick window made of aluminized Mylar, which 
      serves as the cathode and is set at 4260~V.
   \item The drift region is filled with the same Ne-DME gas mixture and extends 
      from the cathode to the first GEM, 60 mm away 
      from the beam axis. The average electric field in this region is about 550~V/cm 
      and perpendicular to the beam.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66 mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower Voltage to
      obtain a strong ($\sim$1600 V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charge
     from the GEMs. Pre-amplifiers are plugged directly on its outer side and
     transmit the signal to the data acquisition electronics.
\end{itemize}

@@

The drift gas used in the experiment is a 80-20\% Ne-DME mixture. This choice 
has been made in order to balance the energy deposit, which is critical
for proper particle identification, with a reasonable
Lorentz angle. Calculations using the MAGBOLTZ program~\cite{MAGBOLTZ} 
showed that with the axial 5~T field, we would have a Lorentz angle of 
about 23$^\circ$ with this gas mixture. This is already a significant angle,
a heavier gas would have larger angle increasing significantly
the length of the drift paths as the electron would 
make a long spiral around the beam line to reach the GEMs.

One of the important challenge in developing the radial TPC was how to obtain a 
good support structure for the GEM foils to allow a tractable 
installation of the GEMs. At the same time, we wanted to 
keep the material budget small in the forward region where we detect other 
particles in subsequent detectors. We successfully realized these prerequisites 
by using fiber glass rings glued to each end of the GEM foils to form self 
supporting cylinders that could be installed independently in the RTPC after 
all gluing and soldering operations. The rigidity of the GEM foils was enough
for the structure to be self-supporting and only the backward part of the 
cylinder was fixed to the main mechanical structure. This design only left a 
light fiberglass ring in the forward region, reducing to a minimum secondary 
interactions.

@@

The RTPC electron collection system has 3200 readout pads. These elements are
parallel to the GEMs, 69 mm from the central axis.
Figure~\ref{fig:PADs} shows a schematic drawing of the size and 
configuration of the pads. Each readout pad is 5 mm in length and 4.45 mm in 
width. The shift between the rows permits the reduction of aliasing. Each half of the 
RTPC has 40 rows and 40 columns of pads. The shaded region in Fig.~\ref{fig:PADs} 
shows how pads are grouped to 16 channel pre-amplifiers. The pre-amplifier boards, 
already employed in the BoNuS RTPC~\cite{BONUS-NIM}, serve the dual purpose of 
inverting the RTPC signals polarity -- from negative to positive -- to match the 
requirements of the subsequent readout system, and driv the 6~m long ribbon 
cable that connects to it.

\begin{figure*}[tb]
   \centering
   \includegraphics[width=.75\textwidth]{fig/ReadoutScheme.png}
   \caption[]{Schematic representation of the RTPC readout system, showing
   the different elements of the front end electronics.} 
   \label{fig:ReadoutScheme}
\end{figure*}

The readout system is based on the Front End Electronic (FEE) boards originally 
developed for the ALICE TPC readout system at CERN~\cite{ALICE-FEE}. Each 
readout channel is made of three main components: a charge sensitive, low 
impedance shaping amplifier, a 10 bit, 25-MHz digitizer, and a digital circuit 
that implements online-processing algorithms (pedestal subtraction, zero 
suppression, and tail cancellation). The amplifier is implemented in a 
fully-analogue ASIC, PASA~\cite{ALICE-PASA}, while the subsequent signal 
digitization and elaboration part is implemented in a digital ASIC, 
ALTRO~\cite{ALICE-ALTRO}. Both ASICs host 16 channels. The full readout 
chain is implemented on a single Front End Card (FEC), hosting 128 channels. 
FECs are hosted in custom crates, mounted close to the detector. A readout 
control unit (RCU) board is used to distribute the trigger signal to FECs 
and for data readout. Each RCU can handle up to 25 FECs with communication 
between them performed trough a 
custom back-plane, implementing a low-Voltage signal bus. The RCU communicates 
with the CLAS data acquisition (DAQ) system 
trough a 200 MB/s optical link, connected to a data acquisition PC hosting 
a ReadOut Receiver Card (RORC) used to interface with 
the CLAS DAQ system. An Ethernet link is also present, for slow-controls and 
monitoring.

The triggering and data-readout scheme is as follows. When a Level-1 (L1) trigger 
is received by a FEC-board, a programmable number of samples $N_0$ is digitized 
and stored in a data memory, and hence processed by the digital circuit. Upon 
arrival of Level-2 trigger, the latest processed samples are saved in a 
multi-event memory, individual to each-channel, ready for readout by the RCU.  
In our setup, the RCU receives the trigger signal from the CLAS Trigger 
Supervisor system, and distributes both L1 and L2 signals to all the FECs in 
the crate, with a definite L1-L2 trigger latency. Such latency is programmable, 
and has to be long enough to accommodate the digitization of the $N_0$ samples 
per channel.  In the current implementation, $N_0=100$ samples/channel are 
acquired for each trigger, operating ALTRO at the reduced sampling frequency of 
$10$~MHz. The L1-L2 latency is thus fixed to 15 $\mu$s.

The trigger signal also initiates RCU readout operation from FEC boards. All 
the measured samples from active channels are reported, together with a channel 
identifier and a time-stamp, to the RCU application, that in turns sends them 
on to the main event builder. During data reconstruction, the acquired samples are 
processed to obtain, for each readout pad, the accumulated charge (ADC) and the 
pulse time (T). Since pulse time was obtained as the time-stamp of the first 
sample above threshold, the resolution is equivalent to the ALTRO sampling 
time of 100~ns.

@@

In order to read all the detector readout pads, four FEC crates are used, 
each equipped with 6 boards, plus a RCU. A schematic of the readout system, 
for a single crate, is reported in Fig.~\ref{fig:ReadoutScheme}. This 
configuration permits to reduce the dead time associated to FEC readout 
operations and to scale linearly with the number of boards in the crate. During the 
2009 run, the system was successfully operated with a DAQ rate of 3.1 kHz, with 
a live time of $70 \%$.

@@

The timing information collected for each signal above threshold
is used to infer the origin of the ionization electrons and 
then the trajectory of the detected particle. Going from a collection of 
times to a momentum measurement based on the curvature of a track 
requires a good knowledge of the drift speed 
and drift paths followed by the electrons released in the gas. The recorded ADCs give 
the deposited energy per unit of length ($\small{\frac{dE}{dX}}$) which, 
together with the momentum calculated from the trajectory, enable the particle 
identification.

In this section we will detail the methods used to calibrate the drift speed,
drift paths and gains of the detector. Drift speed and paths were initially
calculated using the MAGBOLTZ~\cite{MAGBOLTZ} program, then refined using
data to account for variations of the run conditions. We always assume 
cylindrical symmetry in the chamber for the calibration, such that none of
the parameters depend on the azimuthal angle $\phi$. The initial MAGBOLTZ
calibration was improved through several iterations of the
process described below, each iteration increasing the number of events 
properly reconstructed in the RTPC. The figures presented in this section
are the ones obtained while performing the last iteration of this long 
calibration process.

@@

\begin{figure}[tb]
   \centering
   \includegraphics[scale=0.42]{fig_2017/hits_time_profile.pdf}
   \caption[]{Time distribution of the collected hits in one experimental run. Trigger
              time is defined as T=15, the time unit is the ALTRO sampling time of 100~ns.} 
   \label{fig:TDC_profile}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[scale=0.42]{fig_2017/RunNumber_61452_TDCmax_Zslice.pdf}
\caption{Maximum time of collected hits as a function of the track
              position on the $z$-axis for one experimental run. } 
\label{fig:RunNumber_61551_TDCmax_Zslice}
\end{figure}

@@

Because of the magnetic field, the drift paths are not linear in the RTPC. So 
to perform the extraction, we make a first approximation with a linear 
dependence between the radius of emission and the time of detection, and then 
refine our result. As it happens, the curvature is minimal and this process 
converges already on the second iteration.

At the end of the extraction procedure, the azimuthal difference between the detection pad and 
the ionization point ($\Delta\phi$) is extracted as a function of time. 
In Fig.~\ref{fig:DELTA_PHI_TDC}, we show the resulting data points for one 
bin, where the drift path is easily identified and eventually fitted for 
implementation in our reconstruction codes.

@@

\begin{figure*}[tb]
   \centering
   \includegraphics[scale=0.73]{fig_2017/f_dedx_p_exp_1st.pdf}
   \caption{$\small{\frac{dE}{dX}}$ vs. $p/q$ distribution for the left half of 
      the RTPC (left) and for the right half (right). Here, 
      $\small{\frac{dE}{dX}}$ is calculated using the gains of the first 
      method.  The lines are theoretical expectations from Bethe-Bloch formula 
   for $^4$He (red), $^3$He (black), $^3$H (blue), $^2$H (pink) and protons (grey).}
\label{fig:dedx_p_exp_1st}
\end{figure*}

The second method is to compare the experimental ADCs to the energy deposited 
for each pad individually in GEANT4 by similar simulated tracks (using the 
same elastic events as for the drift 
paths calibration). This requires a very good GEANT4 simulation 
including drift paths, but also the spread of the charges along the path
before reaching the pad, so that the simulated hits match the experimental 
ones. Moreover, the simulation has to match the DAQ features 
that can lead to cutting out hits. After setting the simulation properly, we 
compared simulation to experiment on an event by event basis as shown in 
Fig.~\ref{fig:EVENT_adc_tdc}. In this step, the gain for each pad is calculated 
as the ratio of the measured ADCs to the simulated deposited energy.  
Then, these gains are refined using correcting factors obtained from a 
sample of good tracks. For each track, we calculate the corrected 
energy deposit on a pad and compare it to the average
deposit recorded by the other pads. 
Results of this second method are shown in Fig.~\ref{fig:dedx_p_exp_2nd},
from these we concluded that the gains of the 
second method match best the theoretical lines.

\begin{figure}[tb!]
   \centering
\includegraphics[scale=0.42]{fig_2017/Event_ADC_Graph_8799283.pdf}
\caption{Simulated (upper) and experimental (lower) ADC and T distributions 
of a track. The colors indicate the pads, same color in top and bottom indicate 
that they are the same pad.}
\label{fig:EVENT_adc_tdc}
\end{figure}

\begin{figure*}[tb]
\centering
\includegraphics[scale=0.73]{fig_2017/f_dedx_p_exp_2nd.pdf}
\caption{$\small{\frac{dE}{dX}}$ vs. $p/q$ distribution for the left half of 
   the RTPC (left) and for the right half (right). Here, 
   $\small{\frac{dE}{dX}}$ is calculated using the gains of the second method.  
   The lines are theoretical expectations from Bethe-Bloch formula for $^4$He 
   (red), $^3$He (black), $^3$H (blue), $^2$H (pink) and protons (grey).}
\label{fig:dedx_p_exp_2nd}
\end{figure*}

@@

\begin{figure}[tb!]\centering
\includegraphics[scale=0.25]{fig/noisy_pad_before_rejection2.png}
\includegraphics[scale=0.25]{fig/noisy_pad_after_rejection2.png}
\caption{The ADC vs. T spectrum for an example noisy channel before (top) and 
after (bottom) noise rejection algorithms.  Only hits associated with tracks 
are included, and the selection of events and tracks is the same in both 
plots.}
\label{fig:noise}
\end{figure}

@@

\bibitem{CCref}
   G. Adams et al., The CLAS Cerenkov detector, Nucl. Inst. and Meth. A 465, 
   414 (2001).

@@

\bibitem{FX}
   F.X. Girod et al. (CLAS Collaboration), Measurement of Deeply virtual Compton 
   scattering beam-spin asymmetries, Phys.Rev.Lett. 100 (2008) 162002
